A Hedonic Price Analysis of Traffic Noise in the Twin Cities Housing Market

### 2013-08-28 reports/Draft.tex
working on draft

\usepackage{alltt}
\usepackage[round]{natbib}
\usepackage{caption, amsmath, graphicx}

@@

An important first step towards such an analysis is to examine real estate price fluctuations as a function of noise levels using hedonic regression.  Such work has provided important insight on the nature of this relationship.  Thus, Bateman et al. (2001) found that the negative eﬀect of traﬃc noise ranged from 0.08\% to 2.22\% per dB with a mean around 0.55\% per dB. In a review of 65 noise evaluation studies Navrud (2002) found an average decrease of 0.64\% per dB. \textbf{We need to briefly review other relevant papers here.}  However, with the exception of a few studies that have focused on air traffic noise exposure (we should check on this), the relationship between road traffic noise levels and real estate prices has rarely been explored in the United States, even though the former is perhaps the greatest source of noise in residential neighborhoods (XXX).  One major reason for the thinness of the scholarly literature on this topic in the US lies in the difficulty of modeling noise propagation over the landscape.

To properly analyze the spatial association between real estate prices and exposure to traffic noise at a landscape level, it is necessary to create a noise surface map at sufficiently detailed spatial resolution to account for the complex and heterogeneous interaction between the noise source and the resistance of the landscape to noise propagation.  Implementing such a model can be very difficult.  For one thing, the data needed for the model is very extensive and may not even be readily available (e.g., building footprint and height data).  Furthermore, it is computationally very intensive, requiring more computing resources than a desktop. Fortunately, recent developments in GIS and distributive computing have reduced these difficulties, making it much easier to create a noise surface map at landscape level.

In this paper, we seek to contribute to the hedonic literature by examining the relationship between real estate prices and variation in road traffic noise exposure.  We approach this issue by means of an empirical case study of the Twin Cities Metro Region (TCMR), Minnesota.  In particular, we model traffic noise propagation at landscape level using a spatially explicit approach instead of relying on (what?)…

The value of noise pollution is typically estimated through simulated or related markets because no explicit private market for noise abatement exists. \citet{Nelson1982} is perhaps the best known earliest review of 14 housing market studies of the economic impact of traffic noise on property values in Canada and the United States. More recently, a number of reviews have been produced for European government agencies. \citet{Bateman2001} reviewed the literature for the Scottish Government while \citet{Navrud2002} provided a ``state of the art'' review for the European Commission Environment Directorate General.

The Hedonic Price Method is commonly used to value noise by examining real estate price fluctuations as a function of noise levels. Based on the canonical work by \citet{Rosen1974}, researchers typically estimate a hedonic price function specifying the logged sale price of a property as a linear function of the traffic noise exhibited at the house. Such a specification leads to reporting the impact of traffic noise as a percentage decrease in house price per one decibel (dB) increase in traffic noise. According to \citet{Bateman2001}, the negative effect of traffic noise has ranged from 0.08\% to 2.22\% per dB with a mean around 0.55\% per dB. In a review of 65 noise evaluation studies \citet{Navrud2002} found an average decrease of 0.64\% per dB.

\citet{Nelson2008} produced an updated review of the traffic noise literature for the edited volume \underline{Hedonic Methods in Housing Markets Economics}. The chapter describes several areas of continued interest in the empirical literature, most notably: the measure of traffic noise, spatial modeling of the hedonic price function, and housing market dynamics.

\subsection{Measure of Traffic Noise}
I'm going to need some help writing about noise measures. Perhaps Tsegaye can lend a hand?

Noise data is typically not collected for each every house in a hedonic study. For example, \citet{Huang;Palmquist2001} collect noise data using 128 monitoring stations throughout a suburb of Seattle, WA. Such a strategy forces the researchers to interpolate the data between the monitoring stations in order to associate each house with a noise level. While the researchers do not describe how the interpolation was done, they implicitly admit the potential uncertainty of the method by categorizing traffic noise exposure into 2.5 dB wide categories. Other researchers, such as \citet{Theebe2004a} use 5 dB wide bins, and \citet{Nelson2008} even claims that some studies use bins as wide as 10 dB. Such coarse measurements can lead to serious errors in variables and imprecise estimates of the impact of noise.

Other researchers obtain noise levels from external sources without much description of the variable source. For instance, \cite{MarmolejoDuarteCarlos;GonzalezTamez2009} simply list a ``City Council Sonic Map'',  \citet{Blanco2011} cite the Department for Environment Food and Rural Affairs Noise mapping England project. \citet{Andersson2010} state, ``Information about noise levels is from a study on the health effects of traffic noise conducted in Lerum [Sweden] in 2004 (Ohrstrom et al. 2005). Separate noise calculations were made for railway and road noise for all the houses in Lerum.''

Traffic noise at a location can vary significantly depending on changing traffic conditions. Such variation has led to multiple measures of noise. Some measures eliminate the fluctuations in sound and instead represent the equivalent continuous constant sound pressure at a site. Other indices use noise values that are only exceeded a given percentage of time. For instance, \citet{Huang;Palmquist2001} use dB $L_{10}$ as their noise measure, where $L_{10}$ denotes the value that is exceeded only 10 percent of the time. Still others adjust nighttime noise exposure levels to account for potential increased perceived nuisance. \citet{Baranzini2010} even surveyed inhabitants about their perceptions of the level of noise and their annoyance. The work finds similar results for significant levels of measured noise as compared subjective measures of noise.

\emph{What do we have? There is still a lot of work to be done in this area, better understanding the measures others use as well as what we have and how they compare. I'm surprised at how little information people actually provide about their noise data.}

Spatial resolution of the data. \citet{Theebe2004a} had noise levels for square areas of 100 x 100m.

% Beyond the fact that no direct natural economic markets exist in which consumers can reveal their willingness to pay to avoid traffic noise, researchers also must choose how to measure the nuisance itself. Some researchers (see, for instance, \citet{HughesJr.1992}, \citet{Li2012} \citet{Larsen2012}) attempt to measure the negative externality of traffic by correlating traffic volume with house prices. However, traffic may generate other externalities beyond noise (such as local air pollution and reduced safety for pedestrians and drivers) and so these strategies overestimate the size of the noise externality. 
% 
% The value of traffic noise is important in and of itself. Governments routinely erect noise barriers to shield residential areas from traffic noise. (cite some numbers here!). To aid such work, it is important to understand the benefits of these policies. For instance, until 2011, the Minnesota Department of Transportation used a cost effectiveness threshold for noise mitigation projects of \$3,250 per dB reduced per house. Given that some of the benefits of noise reductions are capitalized in house prices, we are curious to see how our estimated magnitudes compare to the state cost effectiveness cutoff.

\subsection{Spatial Analysis}
Advances in Geographic Information System (GIS) technology and digitized datasets have made research easier to conduct over larger geographies, with larger samples, and while explicitly considering the spatial nature of the housing market. Estimation of the hedonic function using Ordinarly Least Squares (OLS) is typically inappropriate because the housing data commonly exhibit spatial autocorrelation. In particular, spatially autocorrelated error terms will lead to inconsistent estimates of OLS regression standard errors and inappropriate measures of statistical inference.

It is now common to use spatial econometric techniques in the hedonic literature, for instance estimating a spatial lag (in which house prices are modeled as a function of neighboring house prices), or a spatial error model (in which the degree of spatial autocorrelation is also estimated). In the noise literature, for example, \citet{Theebe2004a} econometrically estimates a spatial error model, while \citet{Andersson2010} uses a spatial lag model.

One common reason for significant spatially autocorrelated error terms is hedonic function misspecification. A portion of the hedonic regression errors may be explained by omitted spatial variables. Thus, nearby houses will tend to have similar spatial variable values and then similar error terms. Second, model misspecification can occur due to spatial non-stationarity of the hedonic function. That is, housing submarkets exist and the impact of changing housing characteristics will vary across submarkets. Failing to properly account for such heterogeneity will lead to spatially autocorrelated error estimates.

Some traffic noise research simply explicitly assumes that their data comprise a single market. For example, \citet{Kim2007} assumes that by restricting analysis to residential properties surrounding the Inner Circular highway in the central area of Seoul from 2002 to 2004, that the price of traffic noise will be constant. Other research, such as \citet{Theebe2004a} used political boundaries to segment the market into five separate subarkets in the western part of the Netherlands.

Another strategy is to let the data reveal different relationships within smaller areas of the study. For instance, \citet{Day2007} and \citet{MarmolejoDuarteCarlos;GonzalezTamez2009} use semi-parametric techniques to allow the hedonic relationship to smoothly vary over space. We use techniques similar to the Geographically Weighted Regression method described in \citet{MarmolejoDuarteCarlos;GonzalezTamez2009} to allow the hedonic function to vary over space without having a priori conceptions of where the function will be different.

\subsection{Temporal Variation}
Little work has been done estimating the changes in hedonic environmental (dis)amenities like noise over time. An event-study literature exists using temporal variation in noise, especially noise from airports (see, for example, \citet{Cohen2009}, ). However, these articles usually use the change in noise over time to identify the hedonic price of noise, rather than ask whether the hedonic price has changed.

We estimate regression coefficients over time using a moving window of 12 months in the hedonic analysis over the years 2005 - 2010. Such an analysis allows us to see if and when the implicit price of traffic noise changes within our study timeframe. Our results suggest that the percentage change in house price from an extra decibel of traffic noise actually became more negative after the Great Recession than before.

Our work makes three contributions to the field. First, we utilize new estimates of traffic noise that has never before been used in the hedonic literature. The data explicitly model noise propagation as a function of nearby buildings and landcover, as well as present the results at a fine spatial resolution (10m x 10m). Such estimates, combined with sptially explicit modeling techniques allow us to make our second contribution to the literature. We present one of the first estimates of how the impact of traffic noiseeee varies over space. We find that while the impact of other important control variables in the hedonic function vary over space, the impact of traffic noise does not. Lastly, we present the first estimates of how the Great Recession has affected the hedonic price of traffic noise.

\section{Data}
The 2010 US Census lists the population of the Minneapolis and St. Paul Metropolitan area in Minnesota as almost 3 million residents spread over seven counties. This study examines single family residential home transactions in the Census-defined urban areas of three of the seven counties – Dakota, Ramsey and Washington County (see Figure 1). We obtained data from approximately forty thousand sales transactions between 2005 and 2010 (n=42,095) from the 2010 MetroGIS Regional Parcel Dataset published by the Metropolitan Council. In addition to the geographic location and date of the house sale, we collected structural and locational variables commonly used in the hedonic literature. Table \ref{tab:sumStats} provides a brief description of the variables and some basic summary statistics. Figure \ref{fig:overview} shows an overview of the study area as well as the spatial distribution of the house sales prices. Table \ref{tab:cor} shows a simple correlation matrix of the quantitative variables.

@@

 & min & 25\% & 50\% & mean & 75\% & max & $\sigma$ \\ 
  \hline
  Sale Price (thousands \$) & 98.8 & 195.0 & 241.0 & 265.5 & 314.9 & 675.0 & 102.8 \\ 
  House Size (thousands square feet) & 390 & 1158 & 1628 & 1750 & 2188 & 4000 & 704 \\ 
  Lot Size (acres) & 0.02 & 0.17 & 0.25 & 0.25 & 0.31 & 0.60 & 0.11 \\ 
  Owner Occupancy Dummy & 0.0 & 1.0 & 1.0 & 0.8 & 1.0 & 1.0 & 0.4 \\ 
  Year House was Built & 1850 & 1950 & 1973 & 1967 & 1993 & 2010 & 32 \\ 
  Traffic Noise (dB) & 25.1 & 50.5 & 55.3 & 56.4 & 62.2 & 91.5 & 8.8 \\ 
  Median Income in Census Tract (thousands \$) & 13.9 & 54.6 & 69.1 & 72.6 & 89.7 & 143.3 & 23.6 \\ 
  Elementary Test Scores & 335.6 & 359.8 & 365.4 & 364.5 & 370.4 & 551.7 & 10.6 \\ 
  Distance to Central Business District (km) & 1.1 & 6.8 & 13.2 & 14.7 & 21.8 & 37.1 & 8.9 \\ 
  Distance to nearest Park (km) & 0.0 & 1.1 & 2.2 & 2.6 & 3.8 & 9.7 & 1.9 \\ 
  Distance to nearest Lake (km) & 0.0 & 0.4 & 0.8 & 0.9 & 1.3 & 4.4 & 0.7 \\ 
  Distance to nearest Shopping Center (km) & 0.0 & 0.9 & 1.5 & 1.9 & 2.3 & 10.8 & 1.6 \\ 
   \hline
\end{tabular}
}
\end{table}

\begin{table}
\caption{Simple Correlation Matrix of Quantitative Variables}\label{tab:cor}
\includegraphics[width = \textwidth]{../graphs/CorrelationMatrix}
\end{table}

\subsection{Structural Attributes}
According to \cite{Wilhelmsson2000}, the most common structural attributes included in real-estate hedonic pricing studies are living area, number of bathrooms, age, garage and lot size.  Although the 2010 MetroGIS Regional Parcel Dataset includes structural data on living area, age, garage presence, lot size and owner occupancy for every transaction, we have variables for the number of bedrooms, bathrooms, and size of the garage only for those house sales in one county. We feel confident in our results even without these variables for most of our study data because sensitivity analysis conducted in areas with the additional structural variables revealed very similar estimates when the variables were included and excluded. Additionally, other hedonic work has been published using a similar set of explanatory variables in this area, see for instance, \citet{Sander2009b}.

@@

\subsection{Other Locational Attributes}
A common real estate addage states that the three most important things about real estate are: location, location and location. Knowing where each house is located allows us to also construct a vector of other attributes associated with the sales transaction. For instance, using GIS software we are able to calculate the Euclidean distance to numerous points of interest within the dataset, such as the nearest central business district, shopping centers, parks, and lakes.

We also associate each transaction with its elementary school and include the average 3rd grade Minnesota Comprehensive Assessment (MCA) score for the local elementary school during the year of purchase. Test scores were obtained from the Minnesota Department of Education website. The school district and elementary school attendance boundary spatial information was taken from the Minnesota Geospatial Information Office Clearinghouse Data Catalog. Lastly, a variable denoting the median household income for the surrounding census tract is created through the use of TIGER shapefiles from the 2010 Census Bureau and data from the 2010 American Community Survey (ACS).

\subsection{Time}
Given that we have data from before and after the recession of 2008-09, we subset the data in Table \ref{tab:SumStatsTime} by year of sale to look for differences. In addition to the noticeable decline in nominal sales values, there is also a drop in the number of sales across years. For example, in 2010 there were only 4,206 property sales within the dataset, a more than 50\% reduction from 2005. While the mean sale price declined from roughly \$280,000 pre-crash to \$230,000 post-crash, most of the structural and neighborhood variables are relatively consistent across years. However, traffic noise displays a noticeable difference in the pre- and post-crash market transactions (the mean drops from 58 dB to 52 dB).

@@

\subsection{Do the Regression Coefficients Vary over Time}

We know that house prices fell dramatically during and after the Great Recession of 2008-09. We know less about the patterns of change in the hedonic price functions during and after the fall in prices. For instance, did all house prices simply drop (mostly affecting the intercept term of the hedonic function), or did the implicit price of structural characteristics like the size of the house or its lot decrease, or perhaps the value of environmental amenities disproportionately dropped? This section presents some of the first evidence of how the hedonic price function for housing changed over time surrounding the Great Recession of 2008-09. We ask whether the LWR coefficients exhibit a temporal pattern.

Specifically, we create two new variables, ``\texttt{months since September 2008}'' and a dummy variable ``\texttt{post}'' for all sales after September 2008. We then regress our LWR coefficients on these two variables and their interaction term as shown in equation \eqref{eq:LWRinteraction}. 
\begin{equation}\label{eq:LWRinteraction}
\hat{\beta}_{LWR} = \alpha _0 + \alpha _1 \texttt{*month} + \alpha _2  \texttt{*post} + \alpha _3 \texttt{*post*month} + \epsilon
\end{equation}

The results of these regressions will help us understand the patterns of change in the hedonic price of different characteristics. In particular, significant values for $\hat{\alpha}_2$ suggest that there was a structural break in the LWR coefficients post September 2008, and significant values for $\hat{\alpha}_3$ will imply a difference in the temporal trend of the LWR coefficients.

Note that the tables have not yet been formatted.

\subsubsection{Traffic Noise}

We know of no published research that shows how the hedonic price estimates of traffic noise vary over time surrounding the Great Recession. In this section we show that accounting for the month of the house sale significantly predicts to the estimated LWR coefficients for traffic noise. The linear regression results presented in Table \ref{tab:betaMAX} show that there is a significant negative shock to the estimated LWR coefficients for traffic noise after September 2008. Additionally, the coefficients also begin to a significant negative trend.
\begin{table}[h]
\caption{Regression Results: Dependent Variable = Traffic Noise LWR Coefficients}\label{tab:betaMAX}
\begin{verbatim}
(Intercept) -1.781e-03  4.100e-05 -43.436  < 2e-16 ***
month        6.982e-06  2.005e-06   3.482 0.000498 ***
Post        -7.998e-04  7.209e-05 -11.094  < 2e-16 ***
month:Post  -8.868e-05  4.484e-06 -19.776  < 2e-16 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

@@

Table \ref{tab:LWRnonlinear} estimates the predicted LWR noise coefficient as a linear function of the noise level in decibels and whether or not the sale took place before or after September 2008. The table reveals no statistically measurable linear relationship between the noise coefficient for sales before September 2008 and the level of the noise (that is, the marginal effect is no different at low levels of noise than at high levels of noise). The positive and significant interaction term suggests that the impact may be different for high vs. low levels of noise post-September 2008, but the difference in the predicted marginal effects is so small, that it works about to be a roughly \$25 difference for a \$300,000 house at 50 vs. 75 dB.

Figure \ref{fig:betaMAXvCat} visually displays the estimated relationship between the level of traffic noise, time, and the estimated LWR traffic noise coefficients. We categorize the traffic noise data in 5 dB wide bins like \citet{Theebe2004a} and then estimate the mean LWR traffic noise coefficients in each category while also controlling for whether the sales took place pre- or post-September 2008. We see little to suggest any meaningful non-linearities in the effect of traffic noise on house prices, once we have controlled for the timing of the sales.

\begin{figure}
\makebox[\textwidth][c]{\includegraphics[width = 1.1\textwidth]{../graphs/LWRbetaMAXbyCatTime}}
\caption{The mean LWR traffic noise coefficient by 5 decibel category and pre- vs. post-September 2008. The vertical lines denote the standard error of the estimated mean by time and level category. Note that over 99 percent of our data have Noise levels between 35 and 80 dB and no houses were sold post-September 2008 with noise levels above 85 dB.}\label{fig:betaMAXvCat}
\end{figure}

\section{Discussion}

Our work suffers from at least two noticeable weaknesses. First, we are missing some important structural variables that are commonly used in the real estate hedonic literature. To the extent that structural variables like the number of bedrooms, bathrooms, garage size, or construction quality covary with other variables in our dataset, our regression coefficients will suffer from omitted variable bias. We are somewhat comforted, however, because we do have some additional variables for a subset of our data. We have the number of bedrooms, bathrooms, and garage area for our houses located within Dakota County. Our analysis (Models 1, 2, and 3 described earlier) was repeated on this geographic subset of the data and the results were compared to the estimates obtained from the same analysis when these additional structural variables were omitted. We found strikingly similar estimates of the impact of traffic noise.

\begin{table}[ht]
\begin{center}
\caption{Mean and (Standard Deviation) LWR Noise Coefficients in Dakota County both with and without the additional structural variables available only for the county.}\label{tab:Dak}
\begin{tabular}{lll}
 & \multicolumn{2}{c}{Additional variables?} \\
Locally Weighted Regression Model & \multicolumn{1}{c}{no} & \multicolumn{1}{c}{yes} \\ \hline
Model (1) = Structural Variables & -0.00149 & -0.00147 \\ 
   & (0.00226) & (0.00309) \\ 
Model (2) = (1) + Locational Variables & -0.00136 & -0.00140 \\ 
   & (0.00184) & (0.00218) \\ 
Model (3) = (2) + City Fixed Effects & -0.00140 & -0.00137 \\ 
   & (0.00219) & (0.00176) \\ 
\end{tabular}
\end{center}
\end{table}

Table \ref{tab:Dak} shows that the traffic noise coefficients estimated by our LWR models are similar regardless of whether the additional structural variables were included or not. Welch two-sample t-tests fail to reject the null hypothesis of zero means across our three models (comparing LWR estimates with vs. without the additional structural variables included in the model). Paired t-tests find differences for Model 2 and 3, but the estimated differences are zero to four decimal places and in one case the noise impact estimates with the additional variables are slightly larger and in the other case they are slightly smaller. We also conducted simple linear egressions of the noise coefficients without the additional structural variables on the noise coefficients obtained with the additional structural variables. In all three cases the intercept estimates were close zero with slopes almost exactly equal to one and each also had $R^2$ values over 0.7.

The timing of our independent variable collection is also a potential problem. While our house sales data are collected over the course of six years, some of our other variables were collected at specific points in time and assumed to be constant over the study period. In particular, the traffic noise estimates taken from \citet{Nega2012} are for the year 2007. To the extent that traffic flows and composition signicantly changed over time, our traffic noise variables may be inaccurate for those time periods. For instance, the US Department of Transportation reports that total vehicle miles travelled decreased by up to 2 percent year-on-year during the Great Recession. Thus, our noise estimates may be inaccurate for later time periods and this may bias our estimates of the impact of noise on house values. Future work may better estimate the impact of traffic noise over time by generating noise data to reflect changing traffic patterns.

What other weaknesses should we discuss?

We didn't explicitly model spatial autocorrelation in the error terms.

>>>>

\usepackage{alltt}
\usepackage[round]{natbib}
\usepackage{caption, amsmath, graphicx, hyperref}

@@

An important first step towards such an analysis is to examine real estate price fluctuations as a function of noise levels using hedonic regression.  Such work has provided important insight on the nature of this relationship.  Thus, \citet{Bateman2001} found that the negative eﬀect of traﬃc noise ranged from 0.08\% to 2.22\% per dB with a mean around 0.55\% per dB. In a review of 65 noise evaluation studies \citet{Navrud2002} found an average decrease of 0.64\% per dB. \textbf{We need to briefly review other relevant papers here.}  However, with the exception of a few studies that have focused on air traffic noise exposure \textbf{(we should check on this)}, the relationship between road traffic noise levels and real estate prices has rarely been explored in the United States, even though the former is perhaps the greatest source of noise in residential neighborhoods (XXX).  One major reason for the thinness of the scholarly literature on this topic in the US lies in the difficulty of modeling noise propagation over the landscape.

To properly analyze the spatial association between real estate prices and exposure to traffic noise at a landscape level, it is necessary to create a noise surface map at sufficiently detailed spatial resolution to account for the complex and heterogeneous interaction between the noise source and the resistance of the landscape to noise propagation.  Implementing such a model can be very difficult.  For one thing, the data needed for the model is very extensive and may not even be readily available (e.g., building footprint and height data).  Furthermore, it is computationally very intensive. Fortunately, recent developments in Geographic Information Systems and distributive computing have reduced these difficulties, making it much easier to create a noise surface map at landscape level.

In this paper, we seek to contribute to the hedonic literature by examining the relationship between real estate prices and variation in road traffic noise exposure.  We approach this issue by means of an empirical case study of the St.\ Paul, Minnesota urban area.  In particular, we construct a locally flexible model that allows for a non-stationary relationship between house prices and our explanatory variables over geography and time. Such Locally Weighted Regression (LWR) models have become increasingly common in published work and have been shown to be more accurate than spatial autoregressive models with misspecified weights matrices.

Our work makes three contributions to the field. First, we utilize new estimates of traffic noise that has never before been used in the hedonic literature. The data explicitly model noise propagation as a function of nearby buildings and landcover, as well as present the results at a fine spatial resolution (10m x 10m). Such estimates, combined with spatially explicit modeling techniques allow us to make our second contribution to the literature. We present some of the first non-parametric estimates of how the impact of traffic noise spatially varies within an urban area in the United States. We find that while the impact of other important control variables in the hedonic function vary over space, the impact of traffic noise does not. Lastly, we present the first estimates of how the Great Recession has affected the implicit price of traffic noise. Contrary to other published work, our results suggest that consumers' marginal willingness to pay to avoid traffic noise actually increased after the recession.

\section{Data}
The 2010 US Census lists the population of the Twin Cities Metropolitan Region (Minneapolis and St. Paul and their surrounding areas) as almost 3 million residents spread over seven counties. This study examines single family residential home transactions in the Census-defined urban areas of three of the seven counties – Dakota, Ramsey and Washington County (see Figure 1). We obtained data from approximately forty thousand sales transactions between 2005 and 2010 (n=42,095) from the 2010 MetroGIS Regional Parcel Dataset published by the Metropolitan Council. In addition to the geographic location and date of the house sale, we collected and/or constructed structural and locational variables commonly used in the hedonic literature. Table \ref{tab:sumStats} provides a brief description of the variables and some basic summary statistics. Figure \ref{fig:overview} shows an overview of the study area as well as the spatial distribution of the house sales prices. Table \ref{tab:cor} shows a simple correlation matrix of the quantitative variables.

@@

 & min & 25\% & 50\% & mean & 75\% & max & $\sigma$ \\ 
  \hline
  Sale Price (thousands \$) & 98.8 & 195.0 & 241.0 & 265.5 & 314.9 & 675.0 & 102.8 \\ 
  House Size (feet$^2$) & 390 & 1158 & 1628 & 1750 & 2188 & 4000 & 704 \\ 
  Lot Size (acres) & 0.02 & 0.17 & 0.25 & 0.25 & 0.31 & 0.60 & 0.11 \\ 
  Owner Occupancy (0 = no, 1 = yes) & 0.0 & 1.0 & 1.0 & 0.8 & 1.0 & 1.0 & 0.4 \\ 
  Year House was Built & 1850 & 1950 & 1973 & 1967 & 1993 & 2010 & 32 \\ 
  Traffic Noise (dB) & 25.1 & 50.5 & 55.3 & 56.4 & 62.2 & 91.5 & 8.8 \\ 
  Median Income in Census Tract (thousands \$) & 13.9 & 54.6 & 69.1 & 72.6 & 89.7 & 143.3 & 23.6 \\ 
  Elementary School Standardized Test Scores & 336 & 360 & 365 & 365 & 370 & 552 & 11 \\ 
  Distance to Central Business District (km) & 1.1 & 6.8 & 13.2 & 14.7 & 21.8 & 37.1 & 8.9 \\ 
  Distance to nearest Park (km) & 0.0 & 1.1 & 2.2 & 2.6 & 3.8 & 9.7 & 1.9 \\ 
  Distance to nearest Lake (km) & 0.0 & 0.4 & 0.8 & 0.9 & 1.3 & 4.4 & 0.7 \\ 
  Distance to nearest Shopping Center (km) & 0.0 & 0.9 & 1.5 & 1.9 & 2.3 & 10.8 & 1.6 \\ 
   \hline
\end{tabular}
}
\end{table}

\begin{table}
\caption{Pearson Correlation Coefficient Matrix of Quantitative Variables}\label{tab:cor}
\includegraphics[width = \textwidth]{../graphs/CorrelationMatrix}
\end{table}

\subsection{Structural Attributes}
According to \cite{Wilhelmsson2000}, the most common structural attributes included in real-estate hedonic pricing studies are living area, number of bathrooms, age, garage and lot size.  Although the 2010 MetroGIS Regional Parcel Dataset includes structural data on living area, age, garage presence, lot size and owner occupancy for every transaction, we have variables for the number of bedrooms, bathrooms, and size of the garage only for those house sales in one county. We feel confident in our results even without these independent variables for most of our study data because sensitivity analysis conducted in areas with the additional structural variables revealed very similar estimates when the variables were included and excluded. These results are in section \ref{Discussion} of the the paper. Additionally, other hedonic work has been published using a similar set of explanatory variables in this area, see for instance, \citet{Sander2009b}.

@@

\subsection{Other Locational Attributes}
A common real estate addage states that the three most important things about real estate are: location, location and location. Knowing where each house is located allows us to also construct a vector of other attributes associated with the sales transaction. For instance, using GIS software we are able to calculate the Euclidean distance to numerous points of interest within the dataset, such as the nearest central business district, shopping centers, parks, and lakes. Additionally, a variable denoting the median household income for the surrounding census tract is created through the use of TIGER shapefiles from the 2010 Census Bureau and data from the 2010 American Community Survey (ACS). Lastly, we associate each transaction with its elementary school and include the average 3rd grade Minnesota Comprehensive Assessment (MCA) score for the local elementary school during the year of purchase. Test scores were obtained from the Minnesota Department of Education website. The school district and elementary school attendance boundary spatial information was taken from the Minnesota Geospatial Information Office Clearinghouse Data Catalog.\footnote{\url{http://www.mngeo.state.mn.us/chouse/data.html}}

\subsection{Time}
Given that we have data from before and after the recession of 2008-09, we subset the data in Table \ref{tab:SumStatsTime} by year of sale to look for differences. In addition to the noticeable decline in nominal sales values, there is also a drop in the number of sales across years. For example, in 2010 there were only 4,206 property sales within the dataset, a more than 50\% reduction from 2005. While the mean sale price declined from roughly \$280,000 pre-crash to \$230,000 post-crash, most of the structural and neighborhood variables are relatively consistent across years. However, traffic noise is noticeable different in the pre- and post-crash market transactions (the mean drops from 58 dB to 52 dB).

@@

\subsection{Does the Impact of Traffic Noise Vary over Time?}
Little work has been done estimating the changes in hedonic environmental (dis)amenities like noise over time. An event-study literature exists using temporal variation in noise, especially noise from airports (see, for example, \citet{Cohen2009}). However, these articles usually use the change in noise over time to identify the hedonic price of noise, rather than ask whether the hedonic price has changed.

We know of no published research that shows how the hedonic price estimates of traffic noise vary over time surrounding the Great Recession. In this section we seek to compare consumers' willingness to pay for reduced traffic noise exposure before and after the recession. Specifically, we create two new variables, ``\texttt{months since September 2008}'' and a dummy variable, ``\texttt{post},'' for all sales after September 2008. We then regress our LWR coefficients on these two variables and their interaction term as shown in equation \eqref{eq:LWRinteraction}. 
\begin{equation}\label{eq:LWRinteraction}
\hat{\beta}_{LWR} = \alpha _0 + \alpha _1 \texttt{*month} + \alpha _2  \texttt{*post} + \alpha _3 \texttt{*post*month} + \epsilon
\end{equation}

The results of this regression will help us understand the patterns of change (if any) in the marginal willingness to pay to avoid traffic noise. In particular, $\hat{\alpha _1} \neq 0$ suggests a linear temporal trend in the implicit price of traffic noise, $\hat{\alpha}_2 \neq 0$ suggests that there was a structural break in the LWR coefficients post September 2008, and $\hat{\alpha}_3 \neq 0$ implies a difference in the temporal trend of the LWR coefficients before and after September 2008. The linear regression results presented in Table \ref{tab:betaMAX} show that there is a significant negative shock to the estimated LWR coefficients for traffic noise after September 2008. Additionally, the coefficients also begin to show a significant negative monthly trend after September 2008.
\begin{table}[h]
\caption{Regression Results: Dependent Variable = Traffic Noise LWR Coefficients}\label{tab:betaMAX}
\begin{verbatim}
(Intercept) -1.781e-03  4.100e-05 -43.436  < 2e-16 ***
Month        6.982e-06  2.005e-06   3.482 0.000498 ***
Post        -7.998e-04  7.209e-05 -11.094  < 2e-16 ***
Month:Post  -8.868e-05  4.484e-06 -19.776  < 2e-16 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

@@

Table \ref{tab:LWRnonlinear} estimates the predicted LWR noise coefficient as a linear function of the noise level in decibels and whether or not the sale took place before or after September 2008. The table reveals no statistically measurable linear relationship between the noise coefficient for sales before September 2008 and the level of the noise (that is, the marginal effect is no different at low levels of noise than at high levels of noise). The positive and significant interaction term suggests that the impact may be different for high vs. low levels of noise post-September 2008, but the difference in the predicted marginal effects is so small, that it works about to be a difference of roughly \$25 for a \$300,000 house at 50 vs. 75 dB.

Figure \ref{fig:betaMAXvCat} visually displays the estimated relationship between the level of traffic noise, time, and the estimated LWR traffic noise coefficients. We categorize the traffic noise data in 5 dB wide bins like \citet{Theebe2004a} and then estimate the mean LWR traffic noise coefficients in each category while also controlling for whether the sales took place pre- or post-September 2008. We see little to suggest any meaningful non-linearities in the effect of traffic noise on house prices once we have controlled for the timing of the sales.

\begin{figure}
\makebox[\textwidth][c]{\includegraphics[width = 1.1\textwidth]{../graphs/LWRbetaMAXbyCatTime}}
\caption{The mean LWR traffic noise coefficient by 5 decibel category and pre- vs. post-September 2008. The vertical lines denote the standard error of the estimated mean by time and level category. Note that over 99 percent of our data have traffic noise levels between 35 and 80 dB and no houses were sold post-September 2008 with noise levels above 85 dB.}\label{fig:betaMAXvCat}
\end{figure}

\section{Discussion}\label{Discussion}

Our work suffers from at least two noticeable weaknesses. First, we are missing some important structural variables that are commonly used in the real estate hedonic literature. To the extent that structural variables like the number of bedrooms, bathrooms, garage size, or construction quality covary with other variables in our dataset, our regression coefficients will suffer from omitted variable bias.

We are somewhat comforted, however, because we do have some additional variables for a subset of our data and their inclusion does not change our results. We have variables for the number of bedrooms, number of bathrooms, and garage size for houses located within Dakota County. Our analysis (Models 1, 2, and 3 described earlier) was repeated for sales located within Dakota County and the results were compared to the estimates obtained from the same analysis when these additional structural variables were omitted. We found strikingly similar estimates of the impact of traffic noise for these data, regardless of whether or not the additional structural variables were included.

\begin{table}[ht]
\begin{center}
\caption{Mean (Standard Deviation) LWR Traffic Noise Coefficient Estimates in Dakota County with and without Additional Structural Control Variables in Regression.}\label{tab:Dak}
\begin{tabular}{lcc}
 & \multicolumn{2}{p{1.3in}}{Additional structural variables included in regression?} \\
Locally Weighted Regression Model & \multicolumn{1}{c}{no} & \multicolumn{1}{c}{yes} \\ \hline
Model (1) = Structural Variables & -0.00149 & -0.00147 \\ 
   & (0.00226) & (0.00309) \\ 
Model (2) = Model (1) + Locational Variables & -0.00136 & -0.00140 \\ 
   & (0.00184) & (0.00218) \\ 
Model (3) = Model (2) + City Fixed Effects & -0.00140 & -0.00137 \\ 
   & (0.00219) & (0.00176) \\ 
\end{tabular}
\end{center}
\end{table}

Table \ref{tab:Dak} shows that the traffic noise coefficients estimated by our LWR models are similar regardless of whether or not the additional structural variables were included as control variables in the regressions. Welch two-sample t-tests fail to reject the null hypothesis of zero mean differences across our three models (comparing LWR estimates with vs. without the additional structural variables included in the model). Paired t-tests find differences for Model 2 and 3, but the estimated differences are zero to four decimal places and in one case the noise impact estimates with the additional variables are slightly larger and in the other case they are slightly smaller. We also conducted simple linear regressions of the traffic noise coefficients without the additional structural variables on the noise coefficients obtained with the additional structural variables. In all three cases the intercept estimates were close zero and the slope terms almost exactly equal to one. Each regression also had $R^2$ values over 0.7, suggesting that the coefficient estimates obtained with the additional structural explanatory variables are similar to the results obtained without the explanatory variables for Dakota County. We have no reason to believe that our results would be different in the other two counties (Ramsey and Washington) in our study.
